---
title: "Associative Analysis"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Data Analysis Objectives
Perform associative analysis on items sales data to identify relationships between items bought.

### Understanding context
Carrefour Kenya seeks to undertake a project that will inform the marketing department on the most relevant marketing strategies that will result in the highest no. of sales (total price including tax).Associative analysis on sales data would be helpful in identifying customer purchase trends and behaviour.

### Experimental Design

- Problem definition
- Data Loading
- Exploratory data analysis
- Implementation of the solution
- Summary of findings

```{r}
# Load packages
library(arules)
library(arulesViz)
```
```{r}
# Load data
association <- read.transactions("Supermarket_Sales_Dataset 2.csv", sep="," , header = FALSE)

association
```

```{r}
# Preview transactions
inspect(association[1:5])

```
```{r}
# Number of unique items bought
length(itemLabels(association))
```
```{r}
# Get a summary of the transactions
summary(association)
```
```{r}
# Plot the top 10 most frequently purchased items
itemFrequencyPlot(association, topN = 10, type='absolute',col="coral2")
```

Mineral water, eggs and spaghetti are the most frequently purchased items. 



```{r}
# Plot items with a minimum support of 0.1
itemFrequencyPlot(association, support = 0.1, col = "brown3")
```

Seven items have a support of 0.1 or above.


```{r}
# Support per item
support <- sort(itemFrequency(association) * 100)

# Percentage of items with a support less than 0.1%
length(support[support < 0.1])/ 119 * 100

# Percentage of items with a support less than 0.2%
length(support[support < 0.2])/ 119 * 100
```


```{r}
# Build models based on association rules

# Using minimum support of 0.001 and confidence of 0.8
rules_set1 <- apriori(association, parameter = list(supp = 0.001, conf = 0.8, maxlen=10))

# Using minimum support of 0.002 and confidence of 0.6
rules_set2 <- apriori(association, parameter = list(supp = 0.002, conf = 0.6, maxlen=10))

```

```{r}
summary(rules_set1)
```

A minimum support and confidence of 0.001 and .8 respectively generate 74 rules

```{r}
summary(rules_set2)
```
A minimum support and confidence of 0.002 and .6 respectively generate 43 rules

```{r}
# Sort rules based on confidence
inspect(sort(rules_set1, by="confidence", decreasing = T)[1:10])
```
```{r}
# Visualize top 10 rules
top10rules <- head(rules_set1, n = 10, by='confidence')

plot(top10rules, method = "graph",  engine = "htmlwidget")
```
Mushroom cream sauce and pasta are most frequently bought before escalope.


```{r}
inspect(sort(rules_set2, by='confidence', decreasing = T)[1:10])
```

```{r}
top10rules <- head(rules_set2, n=10, by='confidence')
plot(top10rules, method = 'graph', engine = 'htmlwidget')
```


```{r}
# Create shorter association rules
shorter_ruleset <- apriori(association, parameter = list(supp = 0.001, conf = 0.8, maxlen=4))
```
```{r}
inspect(sort(shorter_ruleset, by='confidence', decreasing = T)[1:20])
```

```{r}
top10rules <- head(shorter_ruleset, n=10, by='confidence')
plot(top10rules, method = 'graph', engine = 'htmlwidget')
```
```{r}
# Get association rules associated to shrimp
shrimp_ruleset <- apriori(association, parameter = list(supp=0.001, conf=0.8), appearance = list(default="lhs", rhs="shrimp"))
```
```{r}
inspect(shrimp_ruleset)
```
Customers purchase pasta before purchasing shrimp

```{r}
# Get association rules associated to spaghetti
spaghetti_ruleset <- apriori(association, parameter = list(supp=0.001, conf=0.8), appearance = list(default="lhs", rhs="spaghetti"))


plot(spaghetti_ruleset, method = 'graph', engine = 'htmlwidget')
```
Items most bought before spaghetti are shrimp, salmon and ground beef.

### Summary of findings


1. Mushroom cream sauce and pasta are most often bought together with escalope. Stocking pasta and mushroom cream sauce in the same section or next to each other could increase sales.
2. Shrimp, salmon and ground beef are most often bought before spaghetti, locating the meats section next to the pasta/spaghetti section could boost sales.
3. Package deals of mushroom sauce and spaghetti together.